---
title: "JAGS Notes"
output: html_notebook
---

## Libraries
```{r libraries, message=FALSE}
library(rjags)
library(runjags)
source("2e/DBDA2E-utilities.R")
library(bayesplot)
library(ggplot2)
library(bayestestR)
```


## Model Background

These notes cover how to implement a bayesian model in the ``rjags`` package. The examples for these notes come from chapter 8 of [John Kruschke's book](https://sites.google.com/site/doingbayesiandataanalysis/). 

### Bayes' Formula
As a reminder we will be using ``JAGS`` to solve a model of the form

$$
p(\theta | D) = \frac{p(D|\theta)p(\theta)}{p(D)}
$$
where $p(\theta|D)$ is the posterior distribution, $p(D|\theta)$ is the likelihood distribution, $p(\theta)$ is the prior distribution, and $p(D)$ is a normalizing constant.

In the end this formula can be simplified to $$p(\theta|D) \propto p(D|\theta)p(\theta)$$.

### Fair Coin
For this problem, we will determine the bias of a coin based on collected coin flip data. To accomplish this in ``rjags``, you will need to:

1. Load the data
2. Generate the model
3. Initialize the Markov chains
4. Generate the Markov chains
5. Inspect the Markov chains

#### Load the data

```{r coin-load-data}
data <- read.csv("2e/z15N50.csv")
y <- data$y
Ntotal <- length(y)
dataList <- list(
  "y" = y,
  "Ntotal" = Ntotal
)
```


#### Generate the model

```{r coin-generate-model}
modelString <- "
  model {
    for(i in 1:Ntotal) {
      y[i] ~ dbern(theta)
    }
    theta ~ dbeta(1, 1)
  }
"

#writeLines(modelString, con = "jags-models/coin-model.txt")
```


#### Initialize the chains

``JAGS`` can initalize these values by default, but for large and complex models it is more beneficial to provide a reasonable starting point. For this example, we will use a function that takes a randomized sample of the data and creates a starting point that is approximately equal to the maximum likelihood estimate (MLE).

```{r coin-model-init}

initsList <- function() {
  resampledY <- sample(y, replace = TRUE)
  thetaInit = sum(resampledY) / length(resampledY)
  thetaInit = 0.001 + 0.998*thetaInit
  return (list(theta=thetaInit))
}
```


#### Generate the chains

```{r coin-jags-model}
jagsModel <- jags.model(
  file = "jags-models/coin-model.txt",
  data = dataList,
  inits = initsList(),
  n.chains = 3, #default is 1
  n.adapt = 500 # default is 1000
)
```
##### Burn-in the model

```{r coin-model-burnin}
update(jagsModel, n.iter = 500)
```
##### Generate Samples

```{r coin-generate-samples}
codaSamples <- coda.samples(
  jagsModel,
  variable.names = c("theta"),
  n.iter = 3334
)

```


#### Inspect the chains
```{r coin-model-diagnosis}
plot(codaSamples)

```

##### View the density plot
```{r coin-model-density-plot}
plotPost(codaSamples[, "theta"], main = "Theta", xlab=bquote(theta),
         cenTend = "median", compVal = 0.5, ROPE = c(.45, .55),
         credMass = .90)

```

### Multiple Coins

#### Data
```{r multi-coin-data}
data <- read.csv("2e/z6N8z2N7.csv")
y <- data$y
s <- as.numeric(data$s)
Ntotal <- length(y)
Nsubj <- length(s)
dataList <- list(
  "y" = y,
  "s" = s,
  "Ntotal" = Ntotal,
  "Nsubj" = Nsubj
)

```

#### Model
```{r multi-coin-model}
modelString <- "
  model {
    for(i in 1:Ntotal) {
      y[i] ~ dbern(theta[s])
    }
    for(s in 1:Nsubj){
    theta[s] ~ dbeta(2, 2)
    }
  }
"

#writeLines(modelString, con = "jags-models/two-coins-model.txt")

```

#### Run JAGS
```{r multi-coin-jags}
source("2e/Jags-Ydich-XnomSsubj-MbernBeta.R")
#Generate the chain
mcmcCoda <- genMCMC(data = data, numSavedSteps = 10000)
```

#### Diagnostics
```{r multi-coin-diagnostics}
plot(mcmcCoda)

```

#### Summary
```{r multi-coin-summary}
smryMCMC(mcmcCoda, compVal = NULL, compValDiff = 0.0)

```

#### Plot Posterior
```{r multi-coin-posterior}
plotMCMC(mcmcCoda, data = data, compVal = NULL, compValDiff = 0.0)
```


